Still Missing a Cool Title

### 2006-05-31 talk/dls2006/paper.tex
latexify tables.

The actual transformation from function objects -- i.e. bytecode -- to
flow graph is performed by the Flow Object Space, a short but generic
plug-in component for the Python interpreter of PyPy.  The architecture
of our Python interpreter is shown in figure 4.

\begin{verbatim}
.. topic:: Figure 4 - the interpreter and object spaces

    +------------------------------------------------------+
    |   forest of bytecode objects from the application    |
    +------------------------------------------------------+
    |               Python bytecode interpreter            |
    +--------------------------------+---------------------+
    |      Standard Object Space     |  Flow Object Space  |
    +--------------------------------+---------------------+
\end{verbatim}

@@

The tool-chain has been tested with and can sucessfully apply
transformations enabling various combinations of features. The
translated interpreters are benchmarked using pystone (a [Dhrystone]
derivative traditionally used by the Python community, although it is a
rather poor benchmark) and the classical [Richards] benchmark and
compared against [CPython] 2.4.3 results:

\begin{verbatim}
+------------------------------------+-------------------+-------------------+
|  Interpreter                       | Richards,         | Pystone,          |
|                                    | Time/iteration    | Iterations/second |
+====================================+===================+===================+
|  CPython 2.4.3                     |   789ms    (1.0x) |   40322    (1.0x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c                            |  4269ms    (5.4x) |    7587    (5.3x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-thread                     |  4552ms    (5.8x) |    7122    (5.7x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-stackless                  |  5121ms    (6.5x) |    6060    (6.7x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-gcframework                |  6327ms    (8.0x) |    4960    (8.1x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-stackless-gcframework      |  8743ms   (11.0x) |    3755   (10.7x) |
+------------------------------------+-------------------+-------------------+
|  pypy-llvm-c                       |  3797ms    (4.8x) |    7763    (5.2x) |
+------------------------------------+-------------------+-------------------+
|  pypy-llvm-c-prof                  |  2772ms    (3.5x) |   10245    (3.9x) |
+------------------------------------+-------------------+-------------------+
\end{verbatim}

@@

A complete translation of the pypy-c variant takes about 39 minutes,
divided as follows:

\begin{verbatim}
+-------------------------------------------+------------------------------+
| Step                                      |   Time (minutes:seconds)     |
+===========================================+==============================+
| Front-end                                 |            9:01              |
| (flow graphs and type inference)          |                              |
+-------------------------------------------+------------------------------+
| LLTyper                                   |           10:38              |
| (from RPython-level to C-level graphs     |                              |
| and data)                                 |                              |
+-------------------------------------------+------------------------------+
| Various low-level optimizations           |            6:51              |
| (convert some heap allocations to local   |                              |
| variables, inlining, ...)                 |                              |
+-------------------------------------------+------------------------------+
| Database building                         |            8:39              |
| (this initial back-end step follows all   |                              |
| graphs and prebuilt data structures       |                              |
| recursively, assigns names, and orders    |                              |
| them suitably for code generation)        |                              |
+-------------------------------------------+------------------------------+
| Generating C source                       |            2:25              |
+-------------------------------------------+------------------------------+
| Compiling (\texttt{gcc -O2})              |            3:23              |
+-------------------------------------------+------------------------------+
\end{verbatim}

>>>>

The actual transformation from function objects -- i.e. bytecode -- to
flow graph is performed by the Flow Object Space, a short but generic
plug-in component for the Python interpreter of PyPy.  The architecture
of our Python interpreter is shown in figure \ref{interpobjspace}.

\begin{figure}
  \centering
  \caption{the interpreter and object spaces}
  \label{interpobjspace}
  \begin{tabular}{|c|c|} \hline
\multicolumn{2}{|c|}{forest of bytecode objects from the application}    \\ \hline
\multicolumn{2}{|c|}{Python bytecode interpreter}            \\ \hline
    Standard Object Space     &  Flow Object Space  \\ \hline
  \end{tabular}
\end{figure}

@@

The tool-chain has been tested with and can sucessfully apply
transformations enabling various combinations of features. The
translated interpreters are benchmarked using pystone (a [Dhrystone]
derivative traditionally used by the Python community, although it is
a rather poor benchmark) and the classical [Richards] benchmark and
compared against [CPython] 2.4.3 results and are summarized in table
\ref{perfsumm}.

\begin{table*}
% the layout of this table isn't wonderful, but it's probably OK.
\centering
\caption{Summary of interpreter performance}
\label{perfsumm}
\begin{tabular}{|l|c|c|} \hline
\textbf{Interpreter} & 
\textbf{Richards, Time/iteration} &
\textbf{Pystone, Iterations/second} \\ \hline
CPython 2.4.3                &   789ms  (1.0x) & 40322  (1.0x) \\ \hline
pypy-c                       &  4269ms  (5.4x) &  7587  (5.3x) \\ \hline
pypy-c-thread                &  4552ms  (5.8x) &  7122  (5.7x) \\ \hline
pypy-c-stackless             &  5121ms  (6.5x) &  6060  (6.7x) \\ \hline
pypy-c-gcframework           &  6327ms  (8.0x) &  4960  (8.1x) \\ \hline
pypy-c-stackless-gcframework &  8743ms (11.0x) &  3755 (10.7x) \\ \hline
pypy-llvm-c                  &  3797ms  (4.8x) &  7763  (5.2x) \\ \hline
pypy-llvm-c-prof             &  2772ms  (3.5x) & 10245  (3.9x) \\ \hline
\end{tabular}
\end{table*}

@@

A complete translation of the pypy-c variant takes about 39 minutes,
divided as shown in table \ref{translationtimes}.

\begin{table*}
\centering
\caption{Dividing overall translation time by stage}
\label{translationtimes}
\begin{tabular}{|p{0.65\linewidth}|c|} \hline
\textbf{Step} & \textbf{Time (minutes:seconds)} \\ \hline
Front-end                                 
(flow graphs and type inference)          
& 9:01 \\ \hline
LLTyper                                   
(from RPython-level to C-level graphs     
and data)                                 
& 10:38 \\ \hline
Various low-level optimizations           
(convert some heap allocations to local   
variables, inlining, ...)                 
& 6:51 \\ \hline
Database building                         
(this initial back-end step follows all   
graphs and prebuilt data structures       
recursively, assigns names, and orders    
them suitably for code generation)        
& 8:39 \\ \hline
Generating C source                       
& 2:25 \\ \hline
Compiling (\texttt{gcc -O2})              
& 3:23 \\ \hline
\end{tabular}
\end{table*}
